---
title: "Pipelines"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

<script src="reveal_extra.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code

Lines of code take an **input**, process this with a **function** and then give us an **output**. We have created many variables in going through this process, all with similar names and across a lot of different lines and chunks. Surely theres a more efficient way of doing it so that we are not getting confused with a lot of single-use objects?

## Our code

I've taken all of the code that we used for the task we completed earlier from Prof. DeBruine's data and stuck it all in one chunk, without the `ggplot` part.

```{r message=F}
library(tidyverse)

dat <- read_csv("DeBruine_2005_ProcB.csv")

dat_filtered <- filter(dat, age >= 18)

dat_arranged <- arrange(dat_filtered, age)

dat_selected <- select(dat_arranged, id, sex, age, t_non, t_self, st_non, st_self)

dat_mutated <- mutate(dat_arranged, t_diff = t_self - t_non, st_diff = st_self - st_non)

dat_grouped <- group_by(dat_mutated, sex)

descriptives <- summarise(dat_grouped, 
                          m_age = mean(age), 
                          n = n()) 
```

During this process we created 7 variables, many of which were only used **once**... hardly seems like an efficient use of storage space does it?

Let's see if we can improve matters using **pipes**.

## Our code, condensed

```{r message=F}
library(tidyverse)

dat <- read_csv("DeBruine_2005_ProcB.csv")

dat_mutated_2 <- dat %>%
  filter(age >= 18) %>%
  arrange(age) %>%
  select(id, sex, age, t_non, t_self, st_non, st_self) %>%
  mutate(t_diff = t_self - t_non, 
         st_diff = st_self - st_non)

descriptives_2 <- dat_mutated_2 %>%
  group_by(sex) %>%
  summarise(m_age = mean(age),
            n = n()) 
```

Here is what we have second time around, 3 objects:

- `dat`: the same as before. I like to keep raw data inputs in their own thing.
- `dat_mutated_new`: the combination of all stages from `dat_filtered` up to and including `dat_mutated`
- `descriptives_2`: group_by and summarise, using `dat_mutated_2` as the input

Can you spot the differences between our old code and our new code? 

## What did we even do?

We did a number of things here:

- we used the characters **%>%** at the end of each line
- we stacked multiple lines together, each with **one function**
- the functions **first argument is missing**, which is *always* the data input argument in the tidyverse functions
    - remember we discussed this earlier? 

The **%>%** tells R to take the **output** of the function which preceeds it on that line and put it into the **first argument** of the function that follows on the next line.... just so happens that all of the Tidyverse functions we've used have the **input** as the first argument! Convenient, eh?

When reading off the code, try substituting the **%>%** with the phrase "and then"

```{r eval=FALSE}
dat_mutated_2 <- dat %>% # put dat into dat_mutated_2... AND THEN
  filter(age >= 18) %>% # filter age equal to or greater than 18... AND THEN
  arrange(age) %>% # arrange the result by age ascending... AND THEN
  select(id, sex, age, t_non, t_self, st_non, st_self) %>% # only keep these columns... AND THEN
  mutate(t_diff = t_self - t_non, 
         st_diff = st_self - st_non) # create these new variables
```

## Why Care?

### Code Clarity and Efficiency

There will come a time when you have complex analyses to run, you may have a lot of objects in your environment and not want to retain single use ones which ight just get in your way.

### Nesting Functions

There will also be a time when you want to do more than one thing and, in other programming languages, would have to **nest** functions within functions within functions... and that is literally function-ception. 

If we were to create `dat_mutated_3` to be the same as `dat_mutated_2` by nesting functions inside each other, it would look something like this:

```{r}
dat_mutated_3 <- mutate(
  select(
    arrange(
      filter(dat, age >= 18), age), 
    id, sex, age, t_non, t_self, st_non, st_self), 
  t_diff = t_self - t_non,
  st_diff = st_self - st_non)

```

It will melt your brain, your face and your eyes. Pipes help you voice nesting by taking one output and slotting it straight into another. This makes it **linear, instead of inside out** (where the first thing done is in the centre and you work outwards). This makes it easier for us humans to understand. It also allows you to keep comments about a single function on one line, not spread over at least 2.

Try doing <a href="in_class_cars_task.html">this</a> exercise, then try changing it to piped (using **%>%**) and then try doing it nested.

