---
title: "select()"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

<script src="reveal_extra.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- read_csv("DeBruine_2005_ProcB.csv")
dat_filtered <- filter(dat, age >= 18)
dat_arranged <- arrange(dat_filtered, age)

```

## select(): "selects" columns to keep or discard

This time around we are not going to be using the full data set. This might happen to you a lot if you are using secondary data.

Let's only keep the variables that pertain to the question that we are asking.

Use `select()` to keep the following variables: id, sex, age, t_non, t_self, st_non, st_self

```{r}
dat_selected <- select(dat_arranged, id, sex, age, t_non, t_self, st_non, st_self)
```

Here you can see that the function select takes more arguments than you have seen so far.

- The first is the name of the tibble that we are going to be working from. 
- The rest are the *names of the columns that we want to keep*! **All other columns that go unnamed will be discarded**.

We can also specify the opposite: that is, if we only wanted to discard one column (say `id` just as an example) we could use the code `select(dat_arranged, -id)`. All other columns will be retained in this case, only the named column with the attached minus sign is discarded.

This means that we are saved from typing long lines of code, where mistakes can creep in without us noticing!

### Points to remember

- select acts on *columns*, **not** *rows*
- be careful of your spelling! It is easy to to get a column name wrong here, in which case the code will not produce anything!
    - your career as an R user will be plagued with spelling mistakes... mine certainly has been!

___

Unwanted columns have now been removed, let's calculate our self-preference scores

<center> [mutate(): add columns to a data object](mutate.html) </center>